### 2020-02-24 analysis/paper.Rmd
changing output to word doc

---
title: "Replication in Archaeological Science: Shape as a Measure of Weapon Standardisation"
author: "Team Fresh"
date: "2/24/2020"
output: word_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include=TRUE}
#adding this line as a git test
#install specific versions of pakages called out in the paper
#require(devtools)
library(devtools)

#install_version("geomorph", version = "3.0.6", repos = "http://cran.us.r-project.org")
#install.packages("http://https://cran.r-project.org/src/contrib/Archive/geomorph/geomorph_3.0.6.tar.gz", repos=NULL, type="source")
library(geomorph) # geomorph 3.0.6

#install_version("shapes", version = "1.2.3", repos = "http://cran.rstudio.com/") #"http://cran.us.r-project.org")
library(shapes) # shapes 1.2.3

#not mentioned in the technical appendix, but required in technical code
library(shapes)

#package for lattice replication
library(GGally)

```

```{r set working directory, include=TRUE}
#setwd(dirname(rstudioapi::getSourceEditorContext()$path))
#getwd()
library(here)
```

```{r load data, include=TRUE}
# Read the datafile "lances_M" containing a matrix of dimensional data (metric)

#replication file names did not survive publication!
#lances_M<-read.csv("Data/1-s2.0-S0305440318306757-mmc2.csv",header=TRUE)
lances_M<-read.csv(here("Data","1-s2.0-S0305440318306757-mmc2.csv"),header=TRUE)

# Read the datafile "lances_GMM" containing a matrix of landmark data (geometric morphometric)
#lances_GMM<-read.csv("Data/1-s2.0-S0305440318306757-mmc3.csv",header=TRUE)
lances_GMM<-read.csv(here("Data","1-s2.0-S0305440318306757-mmc3.csv"),header=TRUE)

# Convert the 2D matrix of geometric morphometric data ("lances_GMM") into a 3D array
lances_geo<-arrayspecs(lances_GMM[,2:15],7,2) # 7 landmarks, recording 2 dimensions only (xy coordinates), ignoring the first column of lance IDs

```

```{r analysis 1, include=TRUE}
# Before undertaking any descriptive statistics of the metric dimensions of the lances, it is important to re-sort the data matrix and subset the data.
lances_M$site<-factor(lances_M$site,
                    levels = c("Ejsbol","Nydam", "Illerup"),
                    labels = c("Ejsbol","Nydam", "Illerup"),
                    ordered=T)                                # This simply converts the variable 'site' into an ordered variable of factors with levels.

# The data can be subset into specific features of the lance, so that only cases with complete data are used
blade<-subset(lances_M, point=='yes')                                 # Selects all cases with complete blades
socket<-subset(lances_M, socket=='yes')                               # Selects all cases with complete sockets
total<-subset(lances_M, point=='yes' & socket=='yes')                 # Selects all 'complete' lances with all features

# Pairwise correlations of lance features
cor(socket$socket_length,socket$sock_bdiam,use="pairwise.complete.obs") # 0.52
cor(socket$sock_bdiam,socket$sock_sdiam,use="pairwise.complete.obs")    # 0.52
cor(lances_M$width,lances_M$thickness,use="pairwise.complete.obs")      # 0.53
cor(blade$blade_length,blade$width,use="pairwise.complete.obs")         # 0.58
cor(blade$blade_length,blade$thickness,use="pairwise.complete.obs")     # 0.47

# Principal Components Analysis of Havor lance dimensions
totalpca<-prcomp(~total_length+blade_length+socket_length+sock_bdiam+sock_sdiam+width+thickness,data=total,center=T,scale=T)
totalpca        # By typing 'totalpca' a full summary of the components is printed in the terminal.
str(totalpca)   # More information can be viewed about the PCA results by looking at the 'total' object in more detail.

# Heirarchical Cluster Analysis of Havor lance dimensions
totalhc <-hclust(dist(total[,(c(1,5:11))]), "average")  # The cluster analysis is performed
totalhc1<-as.dendrogram(totalhc)                        # The cluster analysis results are converted into a dendrogram object
plot(totalhc1)                                          # Plots the resulting dendrogram object
```

```{r analysis 2, include=TRUE}

                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
                          #### 4. GEOMETRIC MORPHOMETRICS (GMM): LANDMARKS DATA ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

# Create the links (this specifies the order of landmarks by which to connect linking lines)
outline<-rbind(c(1,2),c(2,3),c(3,4),c(4,5),c(5,6),c(6,7),c(7,1))

                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
                          #### 5. GMM DATA MANIPULATION ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

# Subset the data into different sites
Eco<-lances_geo[,,1:6]    # Ejsbol (n=6)
Nco<-lances_geo[,,7:30]   # Nydam (n=24)
Ico<-lances_geo[,,31:78]  # Illerup (n=48)

```

```{r plotting digitized specs, include=TRUE}
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#
                          #### 6. PLOTTING DIGITISED SPECIMENS ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#

# This function plots the landmarks of specimens recorded (as digitised)
plotAllSpecimens(lances_geo,links=outline)  # Plot landmarks of all specimens (digitised)
plotAllSpecimens(Eco,links=outline)         # Plot landmarks of Ejsbol lances (digitised)
plotAllSpecimens(Nco,links=outline)         # Plot landmarks of Nydam lances (digitised)
plotAllSpecimens(Ico,links=outline)         # Plot landmarks of Illerup lances (digitised)

```

```{r GPA, include=TRUE}

                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
                          #### 7. GENERALISED PROCRUSTES ANALYSIS (GPA) ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

# GPA is performed on the lances, including the semi-landmarks at the socket.
# The 'curves' argument is ignored to have permanent landmarks
lancespr<-gpagen(lances_geo)  # GPA of all lances
Epr<-gpagen(Eco)              # GPA of Ejsbol lances
Npr<-gpagen(Nco)              # GPA of Nydam lances
Ipr<-gpagen(Ico)              # GPA of Illerup lances

```

```{r plotting GPA, include=TRUE}

                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#
                          #### 8. PLOTTING GPA SPECIMENS ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#

# This function plots the landmarks of GPA specimens
plotAllSpecimens(lancespr$coords,links=outline) # Plot landmarks of all specimens (GPA)
plotAllSpecimens(Epr$coords,links=outline)      # Plot landmarks of Ejsbol lances (GPA)
plotAllSpecimens(Npr$coords,links=outline)      # Plot landmarks of Nydam lances (GPA)
plotAllSpecimens(Ipr$coords,links=outline)      # Plot landmarks of Illerup lances (GPA)

# This function plots the shapes of GPA specimens
plotshapes(lancespr$coords,joinline=c(1:7,1),symbol=NA) # Plot shapes of all specimens (GPA)
plotshapes(Epr$coords,joinline=c(1:7,1),symbol=NA)      # Plot shapes of Ejsbol lances (GPA)
plotshapes(Npr$coords,joinline=c(1:7,1),symbol=NA)      # Plot shapes of Nydam lances (GPA)
plotshapes(Ipr$coords,joinline=c(1:7,1),symbol=NA)      # Plot shapes of Illerup lances (GPA)

```

```{r check size vs shape, include=TRUE}
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
                          #### 9. ALLOMETRY: CHECKING IF SHAPE IS RELATED TO SIZE ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

# Manually create a vector containing site
site<-as.factor(c(rep("Ejsbøl",6),rep("Nydam",24),rep("Illerup",48)))
site<-factor(site,levels=c("Ejsbøl","Nydam","Illerup")) # Create a vector for the different sites

# Create a geomorph data frame for analysis
lances_geomorph<-geomorph.data.frame(shape=lancespr$coords,site=site,size=lancespr$Csize,ind=lances_GMM$site_lance.id) # Parse the data into a geomorph data frame

# We can test to see if there is an association between centroid size and shape, before conducting further statistical analyses.

# The results are significant, but they cannot be explained by site difference
#lances_ANOVA<-procD.allometry(shape ~ size, ~site, data=lances_geomorph,method="PredLine") # Optional right-hand formula for the inclusion of groups; "~site" can be removed to perform ANOVA on whole dataset

# lances_ANOVA<-procD.allometry(shape ~ size, data=lances_geomorph,method="PredLine") # As above, but with "~site" removed
#plot(lances_ANOVA)
#print(lances_ANOVA)

#Call:
#  procD.allometry(f1 = shape ~ size, f2 = ~site, data = lances_geomorph,      method = "PredLine") 
#
#Homogeneity of Slopes Test
#Df     RSS       SS      Rsq      F       Z Pr(>F)
#Common Allometry  74 0.76683                                        
#Group Allometries 72 0.74137 0.025466 0.030754 1.2366 0.67153  0.246
#
#The null hypothesis of parallel slopes is supported
#based on a significance criterion of alpha = 0.05 
#
#Based on the results of this test, the following ANOVA table is most appropriate
#
#Type I (Sequential) Sums of Squares and Cross-products
#Randomized Residual Permutation Procedure Used
#1000 Permutations
#ANOVA effect sizes and P-values based on empirical F distributions
#
#Df      SS       MS     Rsq      F        Z Pr(>F)  
#log(size)  1 0.05502 0.055024 0.06645 5.3098  1.91423  0.015 *
#  site       2 0.00619 0.003096 0.00748 0.2987 -0.80096  0.796  
#Residuals 74 0.76683 0.010363 0.92607                         
#Total     77 0.82805                                          
#---
#  Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

```

```{r test mean shape diff, include=TRUE}
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
                          #### 10. TESTING FOR MEAN SHAPE DIFFERENCE: HOTELLING T^2 STATISTIC AND GOODALL F STATISTIC ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

# For testing the difference in mean shape (between sites) we use the raw digitised coordinates.
# By using 'scale=TRUE' in our argument, procrustes (with scaling) is carried out in the procedure.

test1<-testmeanshapes(Eco,Nco,resamples=1000,replace=TRUE,scale=TRUE) # Comparing the mean shape of Ejsbol with Nydam
# Results:
# Hotelling T^2 statistic: F=3.595 P=0.099
# Goodall F statistic: F= 0.084 P= 0.806

test2<-testmeanshapes(Eco,Ico,resamples=1000,replace=TRUE,scale=TRUE) # Comparing the mean shape of Ejsbol with Illerup
# Results:
# Hotelling T^2 statistic: F=5.746 P=0.004
# Goodall F statistic: F= 0.752 P= 0.430

test3<-testmeanshapes(Ico,Nco,resamples=1000,replace=TRUE,scale=TRUE) # Comparing the mean shape of Illerup with Nydam
# Results:
# Hotelling T^2 statistic: F=0.713 P=0.739
# Goodall F statistic: F= 0.829 P= 0.490

```

```{r ANOVA test, include=TRUE}
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#
                          #### 11. PROCRUSTES ANALYSIS OF VARIANCE (ANOVA) ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#

# The Procrustes ANOVA is identical to permutational-MANOVA.
# We want to perform statistical tests for mean shape difference on the GPA specimens.

# Procustres ANOVA
lances_ANOVA2<-procD.lm(shape~site,data=lances_geomorph, iter=99)
print(lances_ANOVA2)

#Call:
#  procD.lm(f1 = shape ~ site, iter = 99, data = lances_geomorph)
#
#Type I (Sequential) Sums of Squares and Cross-products
#Randomized Residual Permutation Procedure Used
#100 Permutations
#ANOVA effect sizes and P-values based on empirical F distributions
#
#Df      SS        MS     Rsq      F        Z Pr(>F)
#site       2 0.01286 0.0064307 0.01553 0.5917 -0.44955    0.7
#Residuals 75 0.81519 0.0108691 0.98447                       
#Total     77 0.82805       
```

```{r PCA, include=TRUE}

                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#
                          #### 12. PRINCIPAL COMPONENTS ANALYSIS (PCA) OF GENERALISED PROCRUSTES ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#

# PCA is carried out on procrustes aligned specimens.
lancespca<-plotTangentSpace(lancespr$coords,label=TRUE)
lancespca       # By typing 'lancespca' a full summary of the components is printed in the terminal.
str(lancespca)  # More information can be viewed about the PCA results by looking at the 'lancespca' object in more detail.
```

```{r JCA, include=TRUE}
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#
                          #### 13. HEIRARCHICAL CLUSTER ANALYSIS (HCA) OF GENERALISED PROCRUSTES ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#

# HCA is carried out on procustes aligned specimens.
hc<-hclust(dist(two.d.array(lancespr$coords)), "average")  # The cluster analysis is performed
hc1<-as.dendrogram(hc)                  # The cluster analysis results are converted into a dendrogram object
plot(hc1)                               # Plots the resulting dendrogram object

```

```{r bilateral symmetry, include=TRUE}

                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
                          #### 14. BILATERAL SYMMETRY INVESTIGATION ----
                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-

# Define a matrix of the paired landmarks (i.e. S1 and S2, N1 and N2, W1 and W2)
lances_land.pairs<-matrix(c(1,7,2,6,3,5),nrow=3,ncol=2,byrow=TRUE)

# Perform an analysis f the directional and fluctuating asymmetry of Havor lances
lance.sym<-bilat.symmetry(A=shape,ind=ind,object.sym=TRUE,land.pairs=lances_land.pairs,data=lances_geomorph,RRPP=FALSE,iter=999)

# View the results (ANOVA table)
lance.sym

#Call:
#  bilat.symmetry(A = shape, ind = ind, object.sym = TRUE, land.pairs = lances_land.pairs,  
#                 data = lances_geomorph, iter = 999, RRPP = FALSE) 
#
#Symmetry (data) type: Object 
#
#Type I (Sequential) Sums of Squares and Cross-products
#Randomization of Raw Values used
#1000 Permutations
#
#Shape ANOVA
#Df      SS        MS     Rsq        F       Z Pr(>F)   
#ind        77 1.55153 0.0201498 0.99820 577.3822 23.2156  0.001 **
#  side        1 0.00011 0.0001059 0.00007   3.0349  1.3652  0.063 . 
#Residuals  77 0.00269 0.0000349 0.00173                           
#Total     155 1.55433                                             
#---
#  Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

# Plot the results
plot(lance.sym, warpgrids=TRUE)
```

The figure that we are attepting to replicate is Figure 8, captioned as a  "Correlation matrix of Havor lance dimensions showing Pearson's r(lowerpanel), accompanying bivariate plots (upperpanel) and  distributions (histograms with kernel density estimation, on the diagonal). Red=Ejsbøl, green=Nydam, blue=Illerup."

Data appear to be drawn from the main lances_M dataframe, but the figure labels include "socket (max)" and "socket (min)" which do not haave obvious analogues in the column names.

A closer review of the paper leads us to the following line: "Although socket 'diameters' were measured, they are referred to here as minimum and maximum socket 'thickness' (p.38." This points us to two fields, which we can test against the summary table that is Table 1:

```{r replication target preparation, include=TRUE, echo=FALSE}

lances_M %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean=mean(sock_sdiam, na.rm = TRUE), n = sum(!is.na(sock_sdiam)))

lances_M %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean=mean(sock_bdiam, na.rm = TRUE), n = sum(!is.na(sock_bdiam)))
```

We can be confident therefore that "sock_bdiam" and "sock_sdiam" are the sources of "socket (max) [thickness (mm)]" and "socket (min) [thickness (mm)]" respectively. Using some custom functions and the excellent ggally package, we can produce a first replication of the figure.

```{r replication figure, include=TRUE, echo=FALSE}

#notes for those who follow:
# check variable names - dont think sock_bdiam etc. is the same as min max width
# color match
# label match

upper_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_smooth(method=loess, fill="None", color="Black", ...) +
    geom_point(aes(color=site, alpha=0.5))
  p
}

diag_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_histogram(bins=10, color="black", fill="cyan", aes(y = ..ncount..)) + 
    geom_density(color="red",aes(y = ..scaled..,color="red"))
  p
}

ggpairs(lances_M, columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() +
        ggsave(here("Images","replication_fig_8v1.png"))

```

The correlation values to do not match the paper. A quick test with the stats passage verifies the results of our plot and confirms that the issue is not a mis-labelling of a spearman coefficient.

```{r replication corr test, include=TRUE, echo=FALSE}
cor(lances_M$socket_length,lances_M$sock_bdiam, use= "pairwise.complete.obs", method = "pearson")
cor(lances_M$socket_length,lances_M$sock_bdiam, use= "pairwise.complete.obs", method = "spearman")

cor(lances_M$sock_sdiam,lances_M$width, use= "pairwise.complete.obs", method = "pearson")
cor(lances_M$sock_sdiam,lances_M$width, use= "pairwise.complete.obs", method = "spearman")
```
```{r replication corr test, include=TRUE, echo=FALSE}
ggpairs(lances_M[complete.cases(lances_M),], columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() +
        ggsave(here("Images","replication_fig_8v2.png"))

```

## Introduction

## Methods

## Results

## Conclusion

## References